### 2021-03-16 quant_estimate/constraints_log_montyhall.tex
WIP : Monty Hall Constraints PDF

\documentclass{article}

\usepackage{algorithm, algpseudocode}
\usepackage{graphicx}

\algnewcommand{\algorithmicforeach}{\textbf{for each}}
\algdef{SE}[FOR]{ForEach}{EndForEach}[1]
{\algorithmicforeach\ #1\ \algorithmicdo}% \ForEach{#1}
{\algorithmicend\ \algorithmicforeach}% \EndForEach
\usepackage{minted}

\title{Monty Hall Constraints}
\author{Sumit Lahiri}

\begin{document}
	\maketitle
	Parse results after processing \texttt{KLEE} dump for \texttt{Monty Hall} Example.
	\begin{minted}{c}
Path : 1
	0 <= door_switch_pse_var_sym
	door_switch_pse_var_sym <= 1
	0 <= choice_pse_var_sym
	choice_pse_var_sym <= 3
	0 <= car_door_sym
	car_door_sym <= 3
	!((!(car_door_sym == 1) && !(choice_pse_var_sym == 1)))
	!((!(car_door_sym == 2) && !(choice_pse_var_sym == 2)))
	!(!(door_switch_pse_var_sym == 0))

Path : 2
	0 <= door_switch_pse_var_sym
	door_switch_pse_var_sym <= 1
	0 <= choice_pse_var_sym
	choice_pse_var_sym <= 3
	0 <= car_door_sym
	car_door_sym <= 3
	!((!(car_door_sym == 1) && !(choice_pse_var_sym == 1)))
	(!(car_door_sym == 2) && !(choice_pse_var_sym == 2))
	!(door_switch_pse_var_sym == 0)
	!(choice_pse_var_sym == 1)

Path : 3
	0 <= door_switch_pse_var_sym
	door_switch_pse_var_sym <= 1
	0 <= choice_pse_var_sym
	choice_pse_var_sym <= 3
	0 <= car_door_sym
	car_door_sym <= 3
	!((!(car_door_sym == 1) && !(choice_pse_var_sym == 1)))
	!((!(car_door_sym == 2) && !(choice_pse_var_sym == 2)))
	!(door_switch_pse_var_sym == 0)
	choice_pse_var_sym == 1

Path : 4
	0 <= door_switch_pse_var_sym
	door_switch_pse_var_sym <= 1
	0 <= choice_pse_var_sym
	choice_pse_var_sym <= 3
	0 <= car_door_sym
	car_door_sym <= 3
	!((!(car_door_sym == 1) && !(choice_pse_var_sym == 1)))
	!((!(car_door_sym == 2) && !(choice_pse_var_sym == 2)))
	!(door_switch_pse_var_sym == 0)
	!(choice_pse_var_sym == 1)

Path : 5
	0 <= door_switch_pse_var_sym
	door_switch_pse_var_sym <= 1
	0 <= choice_pse_var_sym
	choice_pse_var_sym <= 3
	0 <= car_door_sym
	car_door_sym <= 3
	(!(car_door_sym == 1) && !(choice_pse_var_sym == 1))
	!(!(door_switch_pse_var_sym == 0))
	choice_pse_var_sym == car_door_sym

Path : 6
	0 <= door_switch_pse_var_sym
	door_switch_pse_var_sym <= 1
	0 <= choice_pse_var_sym
	choice_pse_var_sym <= 3
	0 <= car_door_sym
	car_door_sym <= 3
	(!(car_door_sym == 1) && !(choice_pse_var_sym == 1))
	!(!(door_switch_pse_var_sym == 0))
	!(choice_pse_var_sym == car_door_sym)

Path : 7
	0 <= door_switch_pse_var_sym
	door_switch_pse_var_sym <= 1
	0 <= choice_pse_var_sym
	choice_pse_var_sym <= 3
	0 <= car_door_sym
	car_door_sym <= 3
	!((!(car_door_sym == 1) && !(choice_pse_var_sym == 1)))
	(!(car_door_sym == 2) && !(choice_pse_var_sym == 2))
	!(!(door_switch_pse_var_sym == 0))
	choice_pse_var_sym == car_door_sym

Path : 8
	0 <= door_switch_pse_var_sym
	door_switch_pse_var_sym <= 1
	0 <= choice_pse_var_sym
	choice_pse_var_sym <= 3
	0 <= car_door_sym
	car_door_sym <= 3
	!((!(car_door_sym == 1) && !(choice_pse_var_sym == 1)))
	(!(car_door_sym == 2) && !(choice_pse_var_sym == 2))
	!(!(door_switch_pse_var_sym == 0))
	!(choice_pse_var_sym == car_door_sym)

Path : 9
	0 <= door_switch_pse_var_sym
	door_switch_pse_var_sym <= 1
	0 <= choice_pse_var_sym
	choice_pse_var_sym <= 3
	0 <= car_door_sym
	car_door_sym <= 3
	(!(car_door_sym == 1) && !(choice_pse_var_sym == 1))
	!(door_switch_pse_var_sym == 0)
	!(choice_pse_var_sym == 2)
	car_door_sym == 2

Path : 10
	0 <= door_switch_pse_var_sym
	door_switch_pse_var_sym <= 1
	0 <= choice_pse_var_sym
	choice_pse_var_sym <= 3
	0 <= car_door_sym
	car_door_sym <= 3
	(!(car_door_sym == 1) && !(choice_pse_var_sym == 1))
	!(door_switch_pse_var_sym == 0)
	!(choice_pse_var_sym == 2)
	!(car_door_sym == 2)

Path : 11
	0 <= door_switch_pse_var_sym
	door_switch_pse_var_sym <= 1
	0 <= choice_pse_var_sym
	choice_pse_var_sym <= 3
	0 <= car_door_sym
	car_door_sym <= 3
	(!(car_door_sym == 1) && !(choice_pse_var_sym == 1))
	!(door_switch_pse_var_sym == 0)
	choice_pse_var_sym == 2
	car_door_sym == 3

Path : 12
	0 <= door_switch_pse_var_sym
	door_switch_pse_var_sym <= 1
	0 <= choice_pse_var_sym
	choice_pse_var_sym <= 3
	0 <= car_door_sym
	car_door_sym <= 3
	(!(car_door_sym == 1) && !(choice_pse_var_sym == 1))
	!(door_switch_pse_var_sym == 0)
	choice_pse_var_sym == 2
	!(car_door_sym == 3)

Path : 13
	0 <= door_switch_pse_var_sym
	door_switch_pse_var_sym <= 1
	0 <= choice_pse_var_sym
	choice_pse_var_sym <= 3
	0 <= car_door_sym
	car_door_sym <= 3
	!((!(car_door_sym == 1) && !(choice_pse_var_sym == 1)))
	(!(car_door_sym == 2) && !(choice_pse_var_sym == 2))
	!(door_switch_pse_var_sym == 0)
	choice_pse_var_sym == 1
	car_door_sym == 3

Path : 14
	0 <= door_switch_pse_var_sym
	door_switch_pse_var_sym <= 1
	0 <= choice_pse_var_sym
	choice_pse_var_sym <= 3
	0 <= car_door_sym
	car_door_sym <= 3
	!((!(car_door_sym == 1) && !(choice_pse_var_sym == 1)))
	(!(car_door_sym == 2) && !(choice_pse_var_sym == 2))
	!(door_switch_pse_var_sym == 0)
	choice_pse_var_sym == 1
	!(car_door_sym == 3)
	\end{minted}
\end{document}