### 2021-04-29 shared/analysis.tex
Benchmark Example 7

\documentclass[usenames,dvipsnames,acmsmall]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[nounderscore]{syntax}
\usepackage{bussproofs}
\usepackage{mathtools}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{minted}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{minted}

\begin{document}
	\begin{figure}
		\begin{minted}{python}
	def ex7(inpt):
		x = 0
		n = 10  # N This is a forall.
		y = 10  # N This is a forall.
		prob = inpt[0] # This is a forall. 
		while(n > 0):
			d = bernoulli.rvs(size=1, p=prob)[0]
			if(d): 
				x = x + y # x is probabilistic here
			n = n-1
		return x

	# Expected Value of "x"
	# E[x] = prob * n * y
	assert((P(x - (prob * n * y)) <= 0) <= 0.6)
		\end{minted}
		\caption{Program listing for finding probabilistic \texttt{assert} violation.}
	\end{figure}
	\begin{figure}
	\begin{minted}{python}
		(0.0000001 < prob_sym),
		(1 <= y_sym),
		(prob_sym <= 1),
		(10 <= n_sym)
	\end{minted}
	\caption{Constraints on \texttt{Path-1} where control flow is in the \texttt{while} loop.}
	\end{figure}
	\begin{figure}
		\begin{minted}{python}
		(0.0000001 < prob_sym),
		(1 <= y_sym),
		(prob_sym <= 1),
		n_sym == 0 # Exit from Loop
		\end{minted}
		\caption{Constraints on \texttt{Path-2} where control flow exits the \texttt{while} loop.}
	\end{figure}
\begin{figure}
	\begin{minted}{c}
	double prob;
	int x, n, y;

	// forall variables
	klee_make_symbolic(&y, sizeof(y), "y_sym");
	klee_make_symbolic(&n, sizeof(n), "n_sym");
	make_pse_symbolic(&prob, sizeof(prob), "prob_sym", 0, 1);

	// Some restrictions on the values.
	klee_assume(y >= 0);
	klee_assume(10 <= n);

	// Work around for silent concretization of prob.
	prob = 0.5;

	// Prob Sym Variables.
	make_pse_symbolic(&x, sizeof(x), "x_pse_sym", 0, 2147483647);

	std::default_random_engine generator;
	std::bernoulli_distribution bernoulli_rvs(prob);

	// This is the "inner" loop which affects the final value of "x".
	while (n > 0)
	{
		int d = bernoulli_rvs(generator);
		if (d)
			x = x + y;
		n = n - 1;
	}

	return 0;
	\end{minted}
	\caption{\texttt{KLEE Transformation} of the program for testing the \texttt{assert}. }
\end{figure}

	\begin{figure}
		\begin{minted}{c}
	long double prob;
	int x, n, y, n_start, y_start;
	long long unsigned int win = 0, loop_run = 0, prob_runs = 0;

	scanf("%d", &n);
	scanf("%d", &y);
	scanf("%Lf", &prob);

	n_start = n;
	y_start = y;

	while (termCount--)
	{
		x = 0;
		int n_loop = n; // maintain "n" for re-runs of loop.

		// program execution
		std::default_random_engine generator;
		std::bernoulli_distribution bernoulli_rvs(prob);

		// "n" gets changed, we need to maintain the loop. 
		while (n_loop > 0)
		{
			int d = bernoulli_rvs(generator);
			if (d)
			{
				// This doesn't favours the "win" condition.
				x = x + y;
			}
			n_loop -= 1;

			// Iterations of the inner loop.
			// How many times the probabilistic loop gets executed?
			prob_runs++;
		}

		// Sample for probability. This is the assert condition.
		if ((double)x - (prob * n * y) <= 0)
			win++;

		// No. of times the program gets executed.
		loop_run++;
	}
		\end{minted}
		\caption{Testing \texttt{Translate()} of the program for testing the \texttt{assert}. }
	\end{figure}
	\begin{table}
		\centering
		\caption{Comparing \texttt{prob(p)} value Vs \texttt{assert} status for low values of \texttt{prob(p)} shown for limited iterations.}
		\label{}
		\begin{tabular}{ c | c | c | c | c | c | c } 
			\toprule
			\texttt{prob-value} & \texttt{Program Runs}& \texttt{Loop Runs} & \texttt{Wins} & \texttt{P((x - (prob * n * y)) <= 0)}  & \texttt{n} & \texttt{y} \\ 
			\hline\hline
			0.0000001298 & 100000 & 5600000 & 100000 & 1.000000 & 56 & 21 \\ \hline
			0.0000001596 & 100000 & 4000000 & 100000 & 1.000000 & 40 & 17 \\ \hline
			0.0000003384 & 100000 & 3600000 & 100000 & 1.000000 & 36 & 28 \\ \hline
			0.0000039147 & 100000 & 2800000 & 100000 & 1.000000 & 28 & 18 \\ \hline
			0.0000077294 & 100000 & 2300000 & 100000 & 1.000000 & 23 & 24 \\ \hline
			0.0000153588 & 100000 & 2700000 & 100000 & 1.000000 & 27 & 16 \\ \hline
			0.0000306176 & 100000 & 2100000 & 100000 & 1.000000 & 21 & 15 \\ \hline
			0.0000611352 & 100000 & 2200000 & 100000 & 1.000000 & 22 & 14 \\ \hline
			0.0001221703 & 100000 & 2000000 & 100000 & 1.000000 & 20 & 11 \\ \hline
			0.0002442406 & 100000 & 2600000 & 100000 & 1.000000 & 26 & 10 \\ \hline
			0.0004883812 & 100000 & 1900000 & 100000 & 1.000000 & 19 & 13 \\ \hline
			0.0009766624 & 100000 & 1800000 & 100000 & 1.000000 & 18 & 9 \\ \hline
			0.0019532248 & 100000 & 2500000 & 100000 & 1.000000 & 25 & 12 \\ \hline
			0.0039063496 & 100000 & 2400000 & 100000 & 1.000000 & 24 & 8 \\ \hline
			{\textbf{ \color[HTML]{036400} 0.0078125992}} & {\textbf{\color[HTML]{036400} 100000}} & {\textbf{\color[HTML]{036400} 1700000}} & {\textbf{\color[HTML]{036400} 0}} & {\textbf{\color[HTML]{036400} 0.000000}} & {\textbf{\color[HTML]{036400} 17}} & {\textbf{\color[HTML]{036400} 7}} \\ \hline
			0.0156250984 & 100000 & 1600000 & 0 & 0.000000 & 16 & 5 \\ \hline
			0.0312500969 & 100000 & 1400000 & 0 & 0.000000 & 14 & 6 \\ \hline
			0.0625000937 & 100000 & 1300000 & 0 & 0.000000 & 13 & 4 \\ \hline
			0.2500000750 & 100000 & 1100000 & 0 & 0.000000 & 11 & 2 \\ \hline
			0.5000000500 & 100000 & 1000000 & 0 & 0.000000 & 10 & 1 \\ \hline
			\bottomrule
		\end{tabular}
	\end{table}
	\begin{figure}
\begin{minted}{python}
	(0.0075243180 < prob_sym),
	(1 <= y_sym),
	(prob_sym <= 0.0078125992),
	(10 <= n_sym)
	\end{minted}
	\caption{Narrowing constraints on \texttt{Path-1} where control flow is in the \texttt{while} loop.}
\end{figure}
\begin{table}
	\centering
	\caption{Comparing \texttt{prob(p)} value to cases Vs \texttt{assert} status for low values of \texttt{prob(p)} on narrowing}
	\label{}
	\begin{tabular}{ c | c | c | c | c | c | c } 
		\toprule
		\texttt{prob-value} & \texttt{Program Runs}& \texttt{Loop Runs} & \texttt{Wins} & \texttt{P((x - (prob * n * y)) <= 0)}  & \texttt{n} & \texttt{y} \\ 
		\hline\hline
		0.0068589250 & 100000 & 2500000 & 100000 & 1.000000 & 25 & 12 \\ \hline
		0.0068665544 & 100000 & 1600000 & 100000 & 1.000000 & 16 & 5 \\ \hline
		0.0068970719 & 100000 & 1400000 & 100000 & 1.000000 & 14 & 6 \\ \hline
		0.0069581071 & 100000 & 1300000 & 100000 & 1.000000 & 13 & 4 \\ \hline
		0.0070801774 & 100000 & 1200000 & 100000 & 1.000000 & 12 & 3 \\ \hline
		0.0073243180 & 100000 & 1100000 & 100000 & 1.000000 & 11 & 2 \\ \hline
		{\textbf{ \color[HTML]{036400} 0.0078125992}} & {\textbf{\color[HTML]{036400} 100000}} & {\textbf{\color[HTML]{036400} 1700000}} & {\textbf{\color[HTML]{036400} 0}} & {\textbf{\color[HTML]{036400} 0.000000}} & {\textbf{\color[HTML]{036400} 17}} & {\textbf{\color[HTML]{036400} 7}} \\ \hline
		0.0156250984 & 100000 & 1600000 & 0 & 0.000000 & 16 & 5 \\ \hline
		\bottomrule
	\end{tabular}
\end{table}
\begin{table}
	\centering
	\caption{Comparing \texttt{prob(p)} value to cases Vs \texttt{assert} status for low values of \texttt{prob(p)} on further narrowing}
	\label{}
	\begin{tabular}{ c | c | c | c | c | c | c } 
		\toprule
		\texttt{prob-value} & \texttt{Program Runs}& \texttt{Loop Runs} & \texttt{Wins} & \texttt{P((x - (prob * n * y)) <= 0)}  & \texttt{n} & \texttt{y} \\ 
		\hline\hline
		0.0075603531 & 100000 & 1300000 & 100000 & 1.000000 & 13 & 4 \\ \hline
		0.0075963883 & 100000 & 1200000 & 100000 & 1.000000 & 12 & 3 \\ \hline
		0.0076684586 & 100000 & 1100000 & 100000 & 1.000000 & 11 & 2 \\ \hline
		0.0077243207 & 100000 & 2100000 & 0 & 0.000000 & 21 & 24 \\ \hline
		0.0077243234 & 100000 & 2800000 & 0 & 0.000000 & 28 & 16 \\ \hline
	{\textbf{ \color[HTML]{036400} 0.0078125992}} & {\textbf{\color[HTML]{036400} 100000}} & {\textbf{\color[HTML]{036400} 1700000}} & {\textbf{\color[HTML]{036400} 0}} & {\textbf{\color[HTML]{036400} 0.000000}} & {\textbf{\color[HTML]{036400} 17}} & {\textbf{\color[HTML]{036400} 7}} \\ \hline
		\bottomrule
	\end{tabular}
\end{table}
\end{document}