### 2021-06-30 shared/Paper-Draft/symbolic_execution.tex
WIP : SymbEx Algorithm

		\subsection{Background}
\label{sec:trad_sym_exec}

\label{sec:pse}
In this section, we present our technique for augmenting traditional symbolic execution to support probabilistic programs with discrete sampling instructions. 
% 
We begin with a short review of a traditional symbolic execution algorithm and then discuss how we calculate exact path probabilities.

% Symbolic Execution Algorithm.
\begin{algorithm}
	\caption{Probabilistic Symbolic Execution Algorithm}
	\label{alg:symbex}
	\begin{algorithmic}[1]
		\Function{SymbEx}{$S : Program$}
		\Let{$\varphi_{paths}$}{[]}, {$E_{s}$} $\leftarrow$ {[]}, $\sigma \leftarrow$ {[]}, $\varphi \leftarrow \emptyset$, $P \leftarrow \{ \}$  \algorithmiccomment{Initialization}
		\Let{$I_{0}$}{\textsc{getStartInstruction}($S$)}
		\State {$E_{s} \leftarrow$  \{[$I_{0}$, $\varphi$, $\sigma$, $P$]\}} \algorithmiccomment{Start State [$I_{0}$, $\varphi$, $\sigma$, $P$]}
		\While{$E_{s} \neq \emptyset$}
		\Let{$\Delta_{c}$}{\textsc{GetNext}($E_{s}$)}, $\Delta_{c}$ $\rightarrow$ [$I_{c}$, $\varphi$, $\sigma$, $P$]  \algorithmiccomment{Get Next state from $E_{s}$}
		\Switch{\textsc{instType}($I_{c}$)} \label{line:inst_case}
		\Case{$\mathtt{x} \xleftarrow{\$} d$} \algorithmiccomment{Create \texttt{fresh} Symbolic Variable} \label{line:make_pse}
		\Let{$I_{s}$}{\textsc{getNextInstruction}($I_{c}$)}
		\Let{$\varphi, \sigma, P$}{{\textbf{PSESample}}({$x,  d$, $\varphi$, $\sigma$, $P$})}
		\State {$E_{s} \leftarrow E_{s} \cup$ \{[$I_{c}$, $\varphi$, $\sigma$, $P$]\} }
		\EndCase
		\Case{$\mathtt{x} \leftarrow e$} \algorithmiccomment{Assignment Instruction} \label{line:assign}
		\Let{$I_{s}$}{\textsc{getNextInstruction}($I_{c}$)}
		\Let{$\varphi, \sigma, P$}{{\textbf{PSEAssignment}}({v, e, $\varphi$, $\sigma$, $P$})}
		\State {$E_{s} \leftarrow E_{s} \cup$ \{[$I_{c}$, $\varphi$, $\sigma$, $P$]\} }
		\EndCase
		\Case{\textbf{if} $(c_{sym})$ \textbf{then} $S_{1}$ \textbf{else} $S_{2}$} \algorithmiccomment{Branch Instruction}
		\Let{$((\varphi \wedge c_{sym}, \sigma,P,p_c), (\varphi \wedge \neg c_{sym}, \sigma,P,1-p_c))$}{{\textbf{PSEBranch}}({$c_{sym}$, $\varphi$, $\sigma$, $P$})} \label{line:branch}
		\Let{$I_{t}$}{\textsc{getStartInstruction}($P_{1}$)} \Let{$I_{f}$} {\textsc{getStartInstruction}($P_{2}$)}	
		\State {$E_{s} \leftarrow E_{s} \cup$ \{[$I_{t}$, $\varphi \wedge c_{sym}$, $\sigma$, $P$]\} }
		\State {$E_{s} \leftarrow E_{s} \cup$ \{[$I_{f}$, $\varphi \wedge \neg c_{sym}$, $\sigma$, $P$]\} }
		\EndCase
		\Case{\textbf{HALT}} \algorithmiccomment{Terminate Instruction}
		\Let{$\varphi_{paths}$}{$\varphi_{paths}$.\textsc{Append}($\varphi$)}		
		\EndCase
		\EndSwitch
		\EndWhile \\
		\Return{$\varphi_{paths}$}
		\EndFunction
	\end{algorithmic}
\end{algorithm}

\textit{Symbolic execution} is a program analysis technique where a program is run on \textit{symbolic} inputs and all program operations are replaced with those which manipulate these symbolic variables.
%
We present a high-level description of how \textit{symbolic execution} works in Alg~\ref{alg:symbex}. 
% 
For each path in the program, a list of \textit{constraints} are stored that encodes the whole path as a \textit{logical}
formula.
% 
These constraints are over \textit{program} variables that have been marked \textit{symbolic} by the user and are stored in the \textit{state} data structure as \textit{metadata} during the \textit{dynamic} symbolic execution of the program.
% 
Apart from \textit{path constraints}, the \textit{state} also contains a mapping of \textit{symbolic} variables to expressions that are either  \textit{symbolic} or \textit{concrete} (in the case of concrete execution of the program) and a list of \textit{instructions} currently getting executed as a part of the \textit{state}.	
% 
During symbolic execution, program state is encoded symbolically in two parts: a conjunctive formula, $\varphi$, consisting of the branch conditions which are true for that particular path, called a \textit{path condition}, and a mapping from program variables to symbolic expressions containing constants and \textit{symbolic variables}, $\sigma$.
% 
Each \textit{instruction} in the program is executed and depending upon the \textit{type} of the \textit{instruction} the \textit{state} along with the \textit{symbolic variables} mapping, ($\sigma$ in Alg~\ref{alg:symbex}) are updated as described in Algorithm~\ref{alg:symbex}.
% 
\textit{Symbolic variables} are assigned values by employing an \textsc{SMT Solver} for solving the current set of \textit{constraints} collected so far in the \textit{program path} and also by \textit{concretizing} some of the values that get evaluated as a result of \textit{concrete} execution (e.g. return value from external function calls).
%
When execution reaches an assignment of the form $\mathtt{x} = e$, where $e$ is a constant or program variable, $\sigma[\mathtt{x}]$ is either just the constant, $e$, or the symbolic representation of $e$, namely $\sigma[e]$.
% 
If $e$ is an expression, we recursively convert each program variable in $e$ to its corresponding symbolic expression found in $\sigma$ to construct a new symbolic execution, $e_{sym}$, and set $\sigma[\mathtt{x}] = e_{sym}$.
% 
When execution reaches a branch guarded by the condition $c$, execution proceeds down both branches, one where $\varphi = \varphi \wedge \sigma[c]$, and the other where $\varphi = \varphi \wedge \neg \sigma[c]$.

\subsection{Adding Probabilistic Sampling}
We present the formalism of our approach on \textbf{pWhile}, a core imperative probabilistic programming language, as a model for more general languages:
% 
\[
S := \mathbf{skip} \mid \mathtt{x} \leftarrow e \mid \mathtt{x} \xleftarrow{\$} d \mid \Delta_1 ; \Delta_2 \mid \mathbf{if}~e~\mathbf{then}~\Delta_1~\mathbf{else}~\Delta_2 \mid \mathbf{while}~e~\mathbf{do}~S 
\]

Above, $\mathtt{x}$ is a program variable, $e$ is an expression, and $d$ is a \textit{discrete} distribution expression which denotes which distribution the sample should be drawn from.
% 
We permit distribution expressions to be (optionally) parameterized by program variables.
% 
For example, $\mathsf{UniformInt}(1,\mathtt{x})$ is a uniform distribution which selects at random a value between 1 and $\mathtt{x}$ (inclusive), where $\mathtt{x} \geq 1$.
% 
In order to support sampling instructions we make the following additions to traditional symbolic execution:
% 
\begin{itemize}
	\item \textit{Probabilistic symbolic variables.} We distinguish two types of symbolic variables: \textit{universal} symbolic variables (identical to those in traditional symbolic execution), and \textit{probabilistic} symbolic variables.
	% 
	For each sampling instruction a new probabilistic symbolic variable is created to denote the result of sample.
	\item \textit{Distribution map.} We add a new mapping from probabilistic symbolic variables to distribution expressions, $P$, which tracks the distribution from which a probabilistic symbolic variable was originally sampled from.
	% 
	This is analogous to the symbolic variable map, $\sigma$, except for mapping probabilistic symbolic variables to distributions instead from program variables to symbolic expressions.
	\item \textit{Path probability.} For each path, we adjoin a path probability expression, $p$, which is parameterized by universal symbolic variables.
	% 
	Now each path can be identified by both its path condition and its probability.
\end{itemize}

\begin{algorithm}[H]
	\caption{PSE Assignment Algorithm}
	\label{alg:assign}
	\begin{algorithmic}[1]
		\Function{PSEAssignment}{$\mathtt{x}, e, \varphi, \sigma, P$}
		\Let{$e_{sym}$}{$\sigma[e]$}
		\State{$\sigma[\mathtt{x}] = e_{sym}$}
		\State\Return{$(\varphi,\sigma,P)$}
		\EndFunction
	\end{algorithmic}
\end{algorithm}

\begin{paragraph}{Assignment}
	For assignment statements of the form $\mathtt{x} \leftarrow e$, where $\mathtt{x}$ is a \textit{program} variable and $e$ is an expression, probabilistic symbolic execution proceeds identically to traditional symbolic execution, as detailed in Alg.~\ref{alg:assign}.
	%
	On the \textit{symbolic execution} side, the \textit{symbolic variable} mapping ($\sigma$) is updated with the result of the assignment operation is stored back in the current \textit{state} for further execution as shown in Line~\ref{line:assign} of Alg~\ref{alg:symbex}.
\end{paragraph}

\begin{algorithm}
	\caption{PSE Sampling Algorithm}
	\label{alg:sample}
	\begin{algorithmic}[1]
		\Function{PSESample}{$\mathtt{x}, d, \varphi, \sigma, P$}
		\Let{$\sigma$}{Generate a fresh probabilistic symbolic variable}
		\State{$\sigma[\mathtt{x}] = \sigma$}
		\State{$P[\sigma] = d$}
		\State\Return{$(\varphi,\sigma,P)$}
		\EndFunction
	\end{algorithmic}
\end{algorithm}

\begin{paragraph}{Sampling}
	For sampling statements, $\mathtt{x} \xleftarrow{\$} d$, where $\mathtt{x}$ is a \textit{program} variable and $d$ is a distribution expression, Alg.~\ref{alg:sample} is used.
	% 
	A fresh probabilistic symbolic variable, $\sigma$, is created, $\sigma$ is updated to be $\sigma[\mathtt{x}] = \sigma$, and the original distribution $d$ is recorded in $P$ by setting $P[\sigma] = d$.
	%
	On the \textit{symbolic execution} side, the updated \textit{Distribution Map} ($P$) with the result of sampling operation is stored back in the current \textit{state} for further execution as shown in Line~\ref{line:make_pse} of Alg~\ref{alg:symbex}.  
\end{paragraph}

\begin{algorithm}
	\caption{PSE Branch Algorithm}
	\label{alg:branch}
	\begin{algorithmic}[1]
		\Function{PSEBranch}{$c, \varphi, \sigma, P$}
		\Let{$c_{sym}$}{$\sigma[c]$}
		\Let{$(\sigma_1,\ldots,\sigma_n)$}{$\dom{P}$}
		\Let{$(d_1,\ldots,d_n)$}{$(P[\sigma_1],\ldots,P[\sigma_n])$}
		\Let{$p_c$}{$\frac{\displaystyle\sum_{(v_1,\ldots,v_n) \in \dom{d_1} \times \cdots\times \dom{d_n}} [(c_{sym} \wedge \varphi)\{\sigma_1 \mapsto v_1,\ldots,\sigma_n \mapsto v_n\}]}{\displaystyle\sum_{(v_1,\ldots,v_n) \in \dom{d_1} \times \cdots\times \dom{d_n}}[\varphi\{\sigma_1 \mapsto v_1,\ldots,\sigma_n \mapsto v_n\}]}$}
		\State\Return{$((\varphi \wedge c_{sym}, \sigma,P,p_c), (\varphi \wedge \neg c_{sym}, \sigma,P,1-p_c))$}
		\EndFunction
	\end{algorithmic}
\end{algorithm}

\begin{paragraph}{Branches (Alg.~\ref{alg:branch})}
	Note that with the inclusion of probabilistic symbolic variables we can now either branch on universal or probabilistic symbolic variables (or both).
	% 
	Intuitively, branches are handled much the same in probabilistic symbolic execution as they are in traditional symbolic execution save for one detail: now branches have a probability associated with them.
	% 
	Given a guard expression $c$, how do we compute the probability of $c$ being true?

	To gain intuition, for now just consider the special case where the guard condition only references universal symbolic variables.
	% 
	Let $c_{sym}$ be the equivalent symbolic expression for the guard $c$ and assume that $c_{sym}$ does not reference any probabilistic symbolic variables.
	% 
	Note that which side of the branch is taken is solely determined by the setting of the univeral symbolic variables.
	% 
	Therefore, one side of the branch must have a probability of 1, and the other side, 0.
	% 
	We use Iverson brackets to formalize this idea; the probability of taking the ``true'' branch is $[c_{sym}]$, and the probability of taking the ``false'' branch is $[\neg c_{sym}]$.

	For probabilistic branches, i.e. guards which branch on probabilistic symbolic variables, computing the branch probability is trickier as given a fixed setting of the universal symbolic variables, it is unclear which branch execution will follow as this is dependent upon the sampling results.
	% 
	Without loss of generality, consider a branch of the form $\mathbf{if}~c~\mathbf{then}~\Delta_1~\mathbf{else}~\Delta_2$.
	% 
	As before, we define $c_{sym} = \sigma[c]$, or the symbolic expression representation of the guard expression, $c$, and we want to compute the probability of taking the ``true'' branch and the ``false'' branch, assuming execution has reached the start of the \texttt{if} condition.
	% 
	Since the path condition $\varphi$ records the necessary constraints on the universal and probabilistic symbolic variables which must hold in order to reach this \texttt{if} condition, we can view this probability a \textit{conditional probability}, or the probability that $c_{sym}$ holds given that $\varphi$ is satisfied.
	% 
	In formal notation, we aim to compute $\Pr[c_{sym} \mid \varphi] = \frac{\Pr[c_{sym}\wedge \varphi]}{\Pr[\varphi]}$.

	For now we restrict our view to uniform distributions, although we can support weighted distributions without further problems.
	% 
	Note that each probabilistic symbolic variable, $\sigma$, is mapped to exactly one distribution, $d$, and therefore, $\sigma \in \dom{d}$.
	% 
	So, assuming there are $n$ probabilistic symbolic variables, $\sigma_1,\ldots,\sigma_n$, and so $n$ distributions, $d_1,\ldots,d_n$, the set of all possible values $\sigma_1,\ldots,\sigma_n$ be is $\mathcal{D} = \dom{d_1} \times \cdots \times \dom{d_n}$.
	% 
	We then count the number of elements (or \textit{assignments}) from $\mathcal{D}$ which satisfy $c_{sym} \wedge \varphi$ and $\varphi$, and divide these two quantities as shown on line 5 of Alg.~\ref{alg:branch}.
	% 
	Note that $p_c$ is not necessarily a value, but rather a symbolic expression containing constants and universal symbolic variables.
	% 
	Additionally, we exploit the fact that the sum of the conditional probabilities of the branch outcomes is 1, which allows us to avoid computing the probability of taking the ``false'' branch directly. \\
	% 
	On the \textit{symbolic execution} side, upon reaching the \textit{branch} instruction at Line~\ref{line:branch} in Alg~\ref{alg:symbex}, two new \textit{states} are created ($\Delta_{true}$ \& $\Delta_{false}$) and the constraint encoding the actual \textit{branch condition} ($c_{sym}$) [Line~\ref{line:true_state}] and it's negation ($\neg c_{sym}$)  [Line~\ref{line:false_state}] are appended to them representing the \textit{true} \& \textit{false} side of the \textit{branch}. 
	%
	These states are then added to the \textit{execution} stack containing \textit{states} that need to be explored next. The process continues untill all the \textit{states} in the \textit{execution} stack are explored.
\end{paragraph}

\begin{figure}[h]
	\centering
	\begin{subfigure}{.3\textwidth}
		\centering
		{\small
			\begin{algorithmic}[1]
				\State{$\mathtt{x}\xleftarrow{\$}\mathrm{UniformInt}(1,3)$}
				\State{$\mathtt{y}\xleftarrow{\$}\mathrm{UniformInt}(1,3)$}
				\If{$\mathtt{x} > 1$}
				\If{$\mathtt{x} < \mathtt{y}$}
				\State\Return{$\mathtt{True}$}
				\EndIf
				\Else
				\State\Return{$\mathtt{False}$}
				\EndIf
			\end{algorithmic}
		}
		\caption{Program}
		\label{fig:exam_program}
	\end{subfigure}
	\begin{subfigure}{.69\textwidth}
		\centering
		{\small
			\begin{forest}baseline,for tree=draw,
				[{$\varphi=\top$},align=center, base=bottom
				[{$\varphi = \sigma_1 > 1$}, align=center, base=bottom, edge label={node [midway,above] {$\frac{2}{3}$} }
				[{$\varphi = (\sigma_1 > 1) \wedge (\sigma_1 < \sigma_2)$}, align=center, base=bottom, edge label={node [midway,above] {$\frac{1}{6}$} }
				[{\texttt{True}}]]
				[{$\varphi = (\sigma_1 > 1) \wedge \neg(\sigma_1 < \sigma_2)$}, align=center, base=bottom, edge label={node [midway,above] {$\frac{5}{6}$} }
				[{\texttt{False}}]]]
				[{$\varphi = \neg(\sigma_1 > 1)$}, align=center, base=bottom, edge label={node [midway,above] {$\frac{1}{3}$} }
				[{\texttt{False}}]]]
			\end{forest}
		}
		\caption{Execution Tree}
		\label{fig:exam_tree}

	\end{subfigure}
	\caption{An example program and its symbolic execution tree}
	\label{fig:example}
\end{figure}

\begin{paragraph}{Example}
	Consider the code in Fig.~\ref{fig:exam_program} and suppose we wish to calculate the probability of the program returning \texttt{True}.
	% 
	Following Alg.~\ref{alg:sample} for lines 1,2, we generate fresh probabilistic symbolic variables for $\mathtt{x}$ and $\mathtt{y}$, $\sigma_1$ and $\sigma_2$, respectively.
	% 
	We also store the distributions which $\sigma_1$ and $\sigma_2$ are samples from, namely the discrete uniform distribution $\mathcal{U}\{1,3\}$.
	% 
	In our notation, we say that $\sigma = \{ \mathtt{x} \mapsto \sigma_1, \mathtt{y} \mapsto \sigma_2 \}$ and $P = \{ \sigma_1,\sigma_2 \mapsto \mathcal{U}\{1,3\} \}$.
	% 
	Now following Alg.~\ref{alg:branch} to process line 4 of Fig.~\ref{fig:exam_program}, note that $\mathcal{D} = \{1,2,3\} \times \{1,2,3\}$ and

	\begin{align*}
		p_c = \Pr[\sigma_1 < \sigma_2 \mid \sigma_1 > 1] &= \frac{\Pr[(\sigma_1 < \sigma_2) \wedge (\sigma_1 > 1)]}{\Pr[\sigma_1 > 1]}\\
		&= \frac{\displaystyle\sum_{(v_1,v_2) \in \mathcal{D}} [(\sigma_1 < \sigma_2) \wedge (\sigma_1 > 1)\{\sigma_1 \mapsto v_1, \sigma_2 \mapsto v_2\}]}{\displaystyle\sum_{(v_1,v_2) \in \mathcal{D}} [(\sigma_1 > 1)\{\sigma_1 \mapsto v_1\}]}\\
		&= \frac{1}{6}
	\end{align*}

	This probability means that the probability of taking the ``true'' branch of the inner \texttt{if} condition is only $\frac{1}{6}$, which makes sense as \texttt{x} is restricted to be either $2$ or $3$, but \texttt{y} can be either 1,2, or 3; however, only one combination of \texttt{x} and \texttt{y} will satisfy $\mathtt{x} < \mathtt{y}$, namely $\mathtt{x} = 2$ and $\mathtt{y} = 3$.
\end{paragraph}

We use a robust dynamic symbolic execution engine, \textsc{KLEE} to generate the \textit{path constraints} corresponding to each path that our tool explores. For the purpose of our implementation, we modify \textsc{KLEE} to support creation of \textit{probabilistic symbolic} variables whose values can be sampled from a distribution and process the \textit{path} constraints that are stored in the \textit{state} corresponding to the current execution of the program at each of \textbf{\textit{assignment}} [Algorithm~\ref{alg:symbex}, Line~\ref{line:assign}], \textbf{\textit{branch}} [Algorithm~\ref{alg:symbex}, Line~\ref{line:branch}] \& \textbf{\textit{sampling}} (\texttt{make\_pse\_symbolic()}) statements. [Algorithm~\ref{alg:symbex}, Line~\ref{line:make_pse}].