### 2021-09-04 draft/evaluation.tex
Reviews

\section{Evaluation}
\label{sec:eval}
We use \textsc{KLEE}\cite{cadar2008}, a robust symbolic execution engine, to implement the \textit{probabilistic symbolic} execution algorithm as described in Alg~\ref{alg:symb_ex}. For the purpose of our implementation, we modify \textsc{KLEE} to support creation of \textit{probabilistic symbolic} variables as described in \S~\ref{sec:pse_intro} whose values can be sampled from a distribution ($\delta$) by invoking \textsc{PSESample} as shown in the \textbf{\textit{sampling}} statement [Alg~\ref{alg:symb_ex}, Line~\ref{line:pse_sampling}]. We use the  \textit{path constraints} ($\varphi$)  from the \textit{state} data-structure (maintained by \textsc{KLEE}) at each of the \textbf{\textit{assignment}} [Alg~\ref{alg:symb_ex}, Line~\ref{line:pse_assignment}] and \textbf{\textit{branch}} [Alg~\ref{alg:symb_ex}, Line~\ref{line:symbex_branch}] statements to get the final encoding ($Enc_a, Enc_t$)  which we solve later. [{\color{red} Zach can describe more on how the actual solving and P-map updates happen on the implementation side}]