### 2021-09-04 draft/rebuttal/parts/detailed_responses.tex

\section{Detailed Comments}

\subsection{Reviewer 1}

\subsection{Reviewer 2}

\subsection{Reviewer 3}

While you do state that you made little effort to optimize your symbolic
execution, the "optimization" direction of future work seems to convey that 
you believe it's just a matter of developing heuristics. While I whole-heartedly agree 
that a solid foundation, like the one presented, should be at the core of 
any reasonable symbolic executor, can you give any intuition as to why you 
expect your approach to scale, when the computation of path probabilities 
and solving them with Z3 seems to already dominate the execution time?