### 2021-09-04 draft/related_work.tex

\section{Related Work}
\label{sec:related}

% In this section, we present our work in light of other closely related
% techniques that have been proposed for Probabilistic Symbolic Execution and also
% contrast on where our work differs.

\paragraph*{Probabilistic Symbolic Execution.}
\citet{geldenhuys_2012} was the first to propose a method for probabilistic
symbolic execution. Given a standard, non-probabilistic program, their
technique treats all inputs as drawn from a uniform distribution, and then
computes the probabilities of taking particular program paths by using model
counting. Later works build on this idea for different applications: analyzing
software reliability~\citep{filieri_2013, borges_2014}, quantifying software
changes~\citep{filieri_2015}, generating performance
distribution~\citep{bihuan_2016}, and evaluating worst-case input distributions
in network programs~\citep{p4wn_2021}. More sophisticated schemes apply
volume-bound computations instead of model counting, which is a performance
bottleneck~\citep{sankaranarayanan_2013,DBLP:journals/pacmpl/AlbarghouthiDDN17}.

These methods all focus on probabilistic programs where all the program inputs
are either known constants, or sampled from a known distribution (often the
uniform distribution). In contrast, our technique works for inputs that are
unknown values, and not drawn from a known distribution. Applying existing
methods for probabilistic symbolic execution to our setting would be
prohibitively expensive, since each possible input would need to be considered,
and input may range over a very large domain (e.g., integers ranging from
$-2^{63}$ to $2^{63}-1$).

\paragraph*{Symbolic inference in probabilistic programs.}
\emph{Probabilistic programming languages} are languages enriched with both
\emph{sampling} and \emph{conditioning} operations. The combination of these two
features allows a probabilistic program to encode a complex distribution; many
models of interest in machine learning can be implemented in this way. A recent
line of research considers how to perform \emph{inference} on such programs:
given an assertion $P$, what is the probability that $P$ holds in the
distribution described by the program? Researchers have considered a variety of
approaches, from weighted model counting~\citep{holtzen_2020}, to analyzing
Bayesian networks~\citep{sampson_2014}, to applying computer algebra
systems~\citep{claret_2013, psi_2016, lambda_psi_2020}.

Most existing work on probabilistic programming languages assumes that all
quantities in the program are drawn from known distributions---while we consider
a family of output distributions, probabilistic programming languages typically
consider a single distribution. It would be interesting to extend our work to
handle conditioning, but it would likely be quite challenging since conditioning
can transform our symbolic branch probabilities in highly complex ways.

\paragraph*{Other automated techniques for probabilistic programs.}
Automated verification of probabilistic programs is an active area of research.
We briefly survey some of the most relevant lines of work.

\textsc{AxProf}~\citep{axprof_2019} is a \emph{statistical testing} approach to
analyzing probabilistic programs, which runs the target program multiple times
on concrete inputs in order to estimate probabilities and expected values in the
output distribution. \textsc{AxProf} is highly efficient, and supports programs
with unknown inputs. However, it can only explore a small subset of the input
space and cannot provide logical guarantees over all possible inputs.

\emph{Probabilistic model checking} is a well-studied method for checking
logical formulas on probabilistic transition
systems~\citep{DBLP:conf/icalp/BaierCHKR97, DBLP:reference/mc/BaierAFK18,
kwiatkowska2011prism, dehnert_2017}. However, probabilistic model checking
targets finite state systems, and performance degrades rapidly as the number of
states grows. Our technique applies to infinite state systems; our examples,
which are finite state, are far too large to encode in existing model checking
tools.

\emph{Abstract interpretation} and \emph{algebraic program analysis} methods
have been developed for probabilistic programs~\citep{CousotM12,
DBLP:conf/pldi/WangHR18}. These methods abstract the probabilistic state, losing
information in exchange for better performance.  In contrast, our method
computes path probabilities exactly, in symbolic form.

Finally, there are many \emph{domain-specific} automated analysis for specific
probabilistic properties: termination and resource
analysis~\citep{Chatterjee:2016:AAQ:2837614.2837639, wang_2021,
DBLP:conf/esop/MoosbruggerBKK21}, accuracy~\citep{Chakarov-martingale, SHA18},
reliability~\citep{DBLP:conf/pldi/CarbinKMR12}, differential
privacy~\citep{DBLP:journals/pacmpl/BartheCKS021} and other relational
properties~\citep{AH17, AH18, DBLP:conf/esop/FarinaCG21}, and long-run
properties of probabilistic loops~\citep{DBLP:conf/atva/BartocciKS19,
DBLP:conf/tacas/BartocciKS20}. Our approach aims at a general-purpose analysis.