### 2021-09-07 paper/parts/experiments.tex
Examples

\section{Experimental Evalutaion}
\label{sec:experiments}

\sumit{What experiments and examples to work on?}
\sumit{In 4 pages}

\begin{itemize}
	\item (RQ1) Experiments
	\item (RQ2) Additional Examples
\end{itemize}